---
title: Project title
author: The Scatterplots
output: github_document
---

```{r load-packages, echo = FALSE}
library(tidyverse)
```

```{r load-data, echo = FALSE}
parks <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-22/parks.csv')
```

## Introduction
(1-2 paragraphs)
Brief introduction to the dataset.
You may repeat some of the information about the dataset provided in the introduction to the dataset on the TidyTuesday repository, paraphrasing on your own terms.
Imagine that your project is a standalone document and the grader has no prior knowledge of the dataset.


Our dataset comes from The Trust for Public Land’s ParkScore Index, an annually
released report which ranks the parks of the 100 most populated cities in the
United States according to 4 key metrics: access, investment, amenities, and
acreage. There are `r nrow(parks)` rows and `r ncol(parks)` columns in the data
set ranging from 2012 through 2020.


The COVID-19 Pandemic has laid bare a multitude of inequities that exist in the
United States. While many people were able to flock to their local parks in 
order to get out of the house, many others were not. Our group is interested in
exploring the Parks Access dataset in order to answer questions regarding access
and equity to the nation’s largest cities’ parks. The Citylab article
accompanying the dataset from TidyTuesday discussed a number of reasons for
unequal access to quality parks and we are hoping to explore these in our project.

## What is the relationship between spending per resident and park size in different U.S. regions and/or cities over time?

### Introduction

(1-2 paragraphs)
Introduction to the question and what parts of the dataset are necessary to answer the question. Also discuss why you’re interested in this question.

### Approach

(1-2 paragraphs)
Describe what types of plots you are going to make to address your question.
For each plot, provide a clear explanation as to why this plot (e.g. boxplot, barplot, histogram, etc.) is best for providing the information you are asking about. 
The two plots should be of different types, and at least one of the two plots needs to use either color mapping or facets.

### Analysis

(2-3 code blocks, 2 figures, text/code comments as needed)
In this section, provide the code that generates your plots. Use scale functions to provide nice axis labels and guides.
You are welcome to use theme functions to customize the appearance of your plot, but you are not required to do so.
All plots must be made with ggplot2. Do not use base R or lattice plotting functions.

### Discussion

(1-3 paragraphs)
In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots.
Speculate about why the data looks the way it does.

## How many amenities do parks with the top 10 and bottom 10 rankings in 2020 have and how does this vary based on what proportion of the top 10 and bottom 10 cities’ land is parkland in 2020?

### Introduction

(1-2 paragraphs)
Introduction to the question and what parts of the dataset are necessary to answer the question. Also discuss why you’re interested in this question.

### Approach

(1-2 paragraphs)
Describe what types of plots you are going to make to address your question.
For each plot, provide a clear explanation as to why this plot (e.g. boxplot, barplot, histogram, etc.) is best for providing the information you are asking about. 
The two plots should be of different types, and at least one of the two plots needs to use either color mapping or facets.

### Analysis

(2-3 code blocks, 2 figures, text/code comments as needed)
In this section, provide the code that generates your plots. Use scale functions to provide nice axis labels and guides.
You are welcome to use theme functions to customize the appearance of your plot, but you are not required to do so.
All plots must be made with ggplot2. Do not use base R or lattice plotting functions.

### Discussion

(1-3 paragraphs)
In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots.
Speculate about why the data looks the way it does.

## Presentation

Our presentation can be found [here](presentation/presentation.html).

## Data 

Include a citation for your data here. 
See http://libraryguides.vu.edu.au/c.php?g=386501&p=4347840 for guidance on proper citation for datasets. 
If you got your data off the web, make sure to note the retrieval date.

## References

List any references here. You should, at a minimum, list your data source.
