---
title: Project title
author: The Scatterplots
output: github_document
---

```{r load-packages, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(jsonlite)
library(scales)
library(maps)

#install.packages("cowplot")  # a gganimate dependency
#devtools::install_github('thomasp85/gganimate')
#remotes::install_github("thomasp85/gganimate@v0.1.1")

library(gganimate)

install.packages("gifski")
library(gifski)
library(png)

#Figure Sizing
knitr::opts_chunk$set(fig.width = 10,
                      fig.asp = 0.618,
                      out.width = "90%")
```

```{r load-data, echo = FALSE}
parks <- read.csv("~/R/project-1-the_scatterplots/data/parks.csv")
cities <- fromJSON("~/R/project-1-the_scatterplots/data/cities.json")
```

### Introduction

Our dataset comes from The Trust for Public Land’s ParkScore Index, an annually
released report which ranks the parks of the 100 most populated cities in the
United States according to 4 key metrics: access, investment, amenities, and
acreage. There are `r nrow(parks)` rows and `r ncol(parks)` columns in the data
set ranging from 2012 through 2020.

The COVID-19 Pandemic has laid bare a multitude of inequities that exist in the
United States. While many people were able to flock to their local parks in 
order to get out of the house, many others were not. Our group is interested in
exploring the Parks Access dataset in order to answer questions regarding access
and equity to the nation’s largest cities’ parks. The Citylab article
accompanying the dataset from TidyTuesday discussed a number of reasons for
unequal access to quality parks and we are hoping to explore these in our project.

## What is the relationship between spending per resident and park size in different U.S. regions and/or cities over time?

### Introduction

To answer our first question (above), we used the "parks" dataset and analyzed the relationship between the variables "spend_per_resident_data", "city", "med_park_size_data", and "year". We thought it'd very interesting to look into how the spending on parks varies across different cities and over the course of the last decade. We were also curious and wanted to analyze how spending per resident and park size varies across different regions/cities. Our group felt as though those were four of the most compelling variables in the dataset and we were all unsure of what the relationship would be between spending per resident and park size versus year and then region/city, so we were excited to create visualizations based upon those variables.

### Approach

(1-2 paragraphs)
Describe what types of plots you are going to make to address your question.

For each plot, provide a clear explanation as to why this plot (e.g. boxplot, barplot, histogram, etc.) is best for providing the information you are asking about. 

The two plots should be of different types, and at least one of the two plots needs to use either color mapping or facets.

To answer our first question addressing the relationship between spending per resident and year, we first created a new variable, called "bins", to break down the spending per resident for each city into four quartiles (where the 1st quartile had the least spending per resident and the 4th quartile had the most). This allowed us to analyze how spending trends shifted over time for each quartile. For this visualization, we created a line graph where each line represents a city and the color represents the quartile (or "bin") that the city falls into. The line graph allowed us to illustrate the trends in different cities' spending per resident over time (from 2012-2020) in what we felt was the clearest way to demonstrate that relationship.

# add second graph approach

### Analysis

```{r question 1}
#data wrangling
fivenum(parks$spend_per_resident_data)

# to remove the $ and change from a categorical variable to a numerical variable 
# selected relevant variables, pivot wider to see what cities have data from every year
# drop na's from cities that do not have spending data from every year
# pivot longer to return dataset to a structure that can be plotted on a line plot

parks_q1 <- parks %>%
  select(year, city, spend_per_resident_data) %>% 
  mutate(across(starts_with("spend_per_resident_data"), ~gsub("\\$", "", .) 
                  %>% as.numeric)) %>% 
  pivot_wider(names_from = "year", 
              values_from = "spend_per_resident_data") %>% 
  drop_na() %>% #this is where we lose it 
  pivot_longer(cols = starts_with("20"),
               names_to = "year", 
               values_to = "spend_per_resident") 

#making the year variable numeric so we can join med_park_size_data back 
parks_q1 <- parks_q1 %>% 
  mutate(year = as.numeric(year))
glimpse(parks_q1)

#joining based on city and year to include med_park_size_data in the dataset
parks_q1 <- parks %>% 
  select(city, year, med_park_size_data) %>% 
  right_join(parks_q1, by = c("city","year")) 

#creating quartile bins for spending per resident, ranges based on five number summary
parks_q1 <- parks_q1 %>% 
  arrange(city) %>% 
  mutate(spend_bins = case_when(
    between(spend_per_resident, 0, 59) ~ "1st quartile",
    between(spend_per_resident, 60, 84) ~ "2nd quartile",
    between(spend_per_resident, 85, 131) ~ "3rd quartile",
    TRUE ~ "4th quartile"
  ))

#creating quartile bins for median park size, ranges based on five number summary
  parks_q1 <- parks_q1 %>% 
    mutate(size_bins = case_when(
    between(med_park_size_data, 0, 3.2) ~ "1st quartile",
    between(med_park_size_data, 3.21, 5.0) ~ "2nd quartile",
    between(med_park_size_data, 5.01, 7.7) ~ "3rd quartile",
    TRUE ~ "4th quartile"
  )) 

q1_plot<- ggplot(parks_q1, aes(x = spend_per_resident, y = med_park_size_data, 
                     size = size_bins, color = spend_bins)) + 
        geom_point() +  
        labs(title = "INSERT TITLE HERE",
            subtitle = "Year: {frame_time}",
             x = "Spending per Resident (in USD)", 
             y = "Median Park Size (in acres)", 
             size = "Size Bins", 
            #it would be great if these legends had the actual values of the bins, maybe we change observation names? 
             color = "Spend Bins") +
        scale_x_continuous(breaks = seq(from = 0, to = 400, by = 50)) + 
        scale_y_continuous(breaks = seq(from = 0, to = 20, by = 5)) +
        scale_color_manual(values = c("#997A6D","#3D7799","#C1D5E0","#B1BC6B")) +
        transition_time(as.integer(year), range = c(2012L, 2020L))

animate(q1_plot)
 #TO DO: Update Color Scale, Fix Y Axis, Try to make the plot wider, add title, rename/better name legends 
```

Links: 
https://github.com/thomasp85/gganimate/wiki/Animation-Composition
https://cran.r-project.org/web/packages/gganimate/gganimate.pdf
https://gganimate.com/
https://www.datanovia.com/en/blog/gganimate-how-to-create-plots-with-beautiful-animation-in-r/
http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html
https://ropensci.org/blog/2018/07/23/gifski-release/
https://gif.ski/
https://github.com/r-rust/gifski
https://gganimate.com/articles/gganimate.html#rendering-1

*note to self: what is the relationship between spending per resident and park size in different U.S. cities over time?*

```{r question 1 plot 2}
#data wrangling
```

(2-3 code blocks, 2 figures, text/code comments as needed)
In this section, provide the code that generates your plots. Use scale functions to provide nice axis labels and guides.
You are welcome to use theme functions to customize the appearance of your plot, but you are not required to do so.
All plots must be made with ggplot2. Do not use base R or lattice plotting functions.

### Discussion

(1-3 paragraphs)
In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots.
Speculate about why the data looks the way it does.

## How many amenities do parks with the top 10 and bottom 10 rankings in 2020 have and how does this vary based on what proportion of the top 10 and bottom 10 cities’ land is parkland in 2020?

### Introduction

For our second question, we wanted to look at parks with the top and bottom 10 rankings in 2020 and compare the percentage of their land being parkland. We also wanted to compare the number of amenities each of those parks have, including dog parks, playgrounds, restrooms, etc. To address these questions, we merged the parks and cities datasets by city, and utilized the "city", "longitude", and "latitude" variables from the cities dataset, as well as the "rank" variable (to determine the top and bottom 10 park rankings) and variable for the amenities from the parks dataset. We thought this would be an interesting question to address as we wanted to illustrate and determine which variables were the most meaningful when determining the park rankings. We wanted to compare and interpret whether the top 10 parks had much more amenities than the bottom 10 parks, and also visualize where those parks are geographically in the US and how much of their cities' respective land is parkland.

### Approach

(1-2 paragraphs)
Describe what types of plots you are going to make to address your question.
For each plot, provide a clear explanation as to why this plot (e.g. boxplot, barplot, histogram, etc.) is best for providing the information you are asking about. 
The two plots should be of different types, and at least one of the two plots needs to use either color mapping or facets.



### Analysis

```{r question-2}
### data wrangling

#top/bottom 10 cities
parks_2020 <- parks %>%
  filter(year == 2020,
         rank <= 10 | rank >= 88)

#matching cities dataframe with parks dataframe
cities <- cities %>%
  filter(state != "Maine") %>%
  mutate(city = case_when(city == "Washington" ~ "Washington, D.C.",
                          city == "Charlotte" ~ "Charlotte/Mecklenburg County",
                          TRUE ~ city)) %>%
  select(city, latitude, longitude) %>%
  rbind(tibble(city = c("Arlington, Virginia"),
               latitude = c(38.8816),
               longitude = c(-77.0910)))

#merging cities and parks dataframes
parks_2020_coords <- left_join(parks_2020, cities, by = "city")

#creating an indicator variable for rank
parks_2020_coords <- parks_2020_coords %>%
  mutate(rank_div = ifelse(rank <= 10, "top", "bottom"))

#dodging overlapping points
parks_2020_coords <- parks_2020_coords %>%
  mutate(longitude = case_when(rank == 1 ~ -93.5,
                              rank == 3 ~ -92.6,
                              rank == 2 ~ -76.6,
                              rank == 4 ~ -77.5,
                              rank == 89 ~ -96.7,
                              rank == 94 ~ -97.5,
                              TRUE ~ longitude),
         updown = ifelse(rank %in% c(3, 89, 2), "down", "up"))

### testing out mapping
ggplot() +
  geom_polygon(data = map_data("state"), aes(x = long, y = lat, group = group),
               fill = "white", color = "gray60") +
  geom_point(data = parks_2020_coords,
             aes(x = longitude, y = latitude, color = rank_div,
                 size = as.numeric(str_extract(park_pct_city_data,"\\d+"))/100)) +
  geom_text(data = parks_2020_coords %>% filter(updown == "up"),
            aes(x = longitude, y = latitude, label = paste0("#",rank)),
            size = 3.5, vjust = -.7, family = "bold") +
  geom_text(data = parks_2020_coords %>% filter(updown == "down"),
            aes(x = longitude, y = latitude, label = paste0("#",rank)),
            size = 3.5, vjust = 1.7, family = "bold") +
  scale_size_continuous(labels = scales::percent) +
  scale_color_manual(values = c("#D55E00", "#009E73")) + 
  labs(x = NULL, y = NULL, size = "% of city that\nis parkland",
       title = "Top and bottom 10 city rankings of parks",
       subtitle = "scaled by % of city that is parkland") +
  coord_map() + 
  theme_void() +
  guides(color = "none") + 
  theme(legend.position = c(.92,.3),
        plot.title = element_text(hjust = 0.1),
        plot.subtitle = element_text(hjust = 0.1))
```

```{r question-2-vis-2}
parks_2020_coords <- parks_2020_coords %>%
  mutate(total_amenities = playground_data + restroom_data + basketball_data)


parks_amenities <- parks_2020_coords %>%
  pivot_longer(cols = c(playground_data, restroom_data, basketball_data), names_to = "amenity", values_to = "value")





#parks_2020_coords <- parks_2020_coords %>%
 # mutate(rank_quartile = case_when(rank <= 5 ~ "1st quartile",
  #                        rank > 5 & rank <= 10 ~ "2nd quartile",
   #                       rank >= 88 & rank < 93 ~ "3rd quartile",
    #                      rank >= 93 ~ "4th quartile"))

#ggplot(data = parks_amenities, mapping = aes(x = city, y = total_amenities, fill = total_amenities)) + 
  #geom_bar(stat = "identity") +
  #geom_text(data = parks_2020_coords, aes(label = rank), hjust = -.5, color = "black", family = "bold") +
  #coord_flip() +
  #labs(y = "Total Amenities", x = NULL)

ggplot(data = parks_amenities, mapping = aes(x = reorder(city, -rank))) + 
  geom_bar(stat = "identity", mapping = aes(y = value, fill = amenity)) +
  geom_text(data = parks_2020_coords, mapping = aes(label = paste0("#",rank), y = total_amenities), hjust = -.1, 
            color = "black",
            family = "bold",
            size = 2.5) +
  coord_flip() +
  scale_fill_discrete(labels = c("Basketball Courts", "Playgrounds", "Restrooms")) +
  guides(fill = guide_legend(reverse = TRUE)) +
  labs(title = "Top and bottom 10 city rankings by amenities",
       y = "Total Amenities per 10K  Residents", x = NULL, fill = "Amenities") +
  theme(plot.title = element_text(hjust = 0),
        plot.subtitle = element_text(hjust = 0)) +
  theme_minimal()
```
```



(2-3 code blocks, 2 figures, text/code comments as needed)
In this section, provide the code that generates your plots. Use scale functions to provide nice axis labels and guides.
You are welcome to use theme functions to customize the appearance of your plot, but you are not required to do so.
All plots must be made with ggplot2. Do not use base R or lattice plotting functions.

### Discussion

(1-3 paragraphs)
In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots.
Speculate about why the data looks the way it does.

For the second question, our first visualization 

## Presentation

Our presentation can be found [here](presentation/presentation.html).

## Data 

Include a citation for your data here. 
See http://libraryguides.vu.edu.au/c.php?g=386501&p=4347840 for guidance on proper citation for datasets. 
If you got your data off the web, make sure to note the retrieval date.

## References

List any references here. You should, at a minimum, list your data source.
