---
title: "Quality of U.S. Public Parks"
#subtitle: "Presentation subtitle (if any)"
author: "The Scatterplots <br> Emely Gutierrez, Zach Khazzam, Matthew Ralph, Kathryn Thomas"
institute: "Duke University"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
# Remove any packages from this list that you're not using
library(tidyverse)
library(jsonlite)
library(scales)
library(maps)
library(gganimate)
library(gifski)
library(png)
library(xaringanthemer)
library(ggrepel)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(
  fig.retina = 3, 
  dpi = 300, 
  fig.width = 6, 
  fig.asp = 0.618, 
  out.width = "70%"
  )
```

```{r load-data, include=FALSE}
parks <- read.csv("~/R/project-1-the_scatterplots/data/parks.csv")
cities <- fromJSON("~/R/project-1-the_scatterplots/data/cities.json")
#parks <- read.csv("~/R/STA313/Projects/project-1-the_scatterplots/data/parks.csv")
#cities <- fromJSON("~/R/STA313/Projects/project-1-the_scatterplots/data/cities.json")
```

```{r style-slides, echo = FALSE}
style_xaringan(
  title_slide_text_color = "#DCDCDC",
  title_slide_background_image = "img/background.png",
  inverse_background_color = "#738148",
  text_color = "#315d1b",
  header_color = "#315d1b"
)
```

# Introduction

--

- The Trust for Public Land’s ParkScore Index: an annually released report which ranks the parks of the 100 most populated cities in the United States according to 4 key metrics: 

--
    - access
    - investment
    - amenities
    - acreage

--

- Covid-19 Pandemic emphasized need for public parks and the inequities that exist between cities

---

class: inverse, middle, center

# Question 1

---

### What is the relationship between spending per resident and park size in different U.S. regions and/or cities over time?

--

- Focus in cities that had spending data for each year from 2012 to 2020.

--

- Metrics of interest: spending per resident, median park size, U.S. Cities and their regions.

--

- Types of plots:
    - annotated scatter plot 
    - line chart over time

---

# Question 1

```{r question-1-plot-1, echo = FALSE, warning = FALSE, message = FALSE}
parks_q1 <- parks %>%
  select(year, city, spend_per_resident_data) %>% 
  mutate(across(starts_with("spend_per_resident_data"), ~gsub("\\$", "", .) 
                  %>% as.numeric)) %>% 
  pivot_wider(names_from = "year", 
              values_from = "spend_per_resident_data") %>% 
  drop_na() %>% #this is where we lose it 
  pivot_longer(cols = starts_with("20"),
               names_to = "year", 
               values_to = "spend_per_resident") 

#making the year variable numeric so we can join med_park_size_data back 
parks_q1 <- parks_q1 %>% 
  mutate(year = as.numeric(year))


#joining based on city and year to include med_park_size_data in the dataset
parks_q1 <- parks %>% 
  select(city, year, med_park_size_data) %>% 
  right_join(parks_q1, by = c("city","year")) 

#creating quartile bins for spending per resident, ranges based on five number summary
parks_q1 <- parks_q1 %>% 
  arrange(city) %>% 
  mutate(spending = case_when(
    between(spend_per_resident, 0, 62) ~ "1st quartile",
    between(spend_per_resident, 63, 94) ~ "2nd quartile",
    between(spend_per_resident, 95, 134) ~ "3rd quartile",
    TRUE ~ "4th quartile"
  ))

#creating quartile bins for median park size, ranges based on five number summary
  parks_q1 <- parks_q1 %>% 
    mutate(size = case_when(
    between(med_park_size_data, 0, 3.0) ~ "1st quartile",
    between(med_park_size_data, 3.01, 4.8) ~ "2nd quartile",
    between(med_park_size_data, 4.81, 7.3) ~ "3rd quartile",
    TRUE ~ "4th quartile"
  )) 

### plot of median park size vs spending per resident over time
  
q1_plot<- ggplot(parks_q1, aes(x = spend_per_resident, y = med_park_size_data, 
                               group = city)) + 
        geom_point(aes(size = size, color = spending)) +   
        labs(title = "Median Park Size vs. Spending Per Resident from 2012-2020 in 37 U.S. Cities",
             subtitle = "Year: {frame_time}",
             x = "Spending per Resident (USD)", 
             y = "Median Park Size (acres)", 
             size = "Park Size", 
             caption = "Quartiles for spending are $0-$62, $63-$94, $95-$134, and $135+ for 1st to 4th quartiles, respectively. 
Quartiles for size are 0-3.2 acres, 3.2-5.0 acres, 5.0-7.7 acres, 7.7+ acres for 1st to 4th quartiles respectively.",
             color = "Spending") +
        theme_minimal()+
        theme(plot.caption = element_text(size = 8, hjust = 0), 
              plot.title = element_text(size = 12), 
              legend.key.size = unit(.65, 'cm'), 
              legend.position = c(.9,.6)) +
        scale_x_continuous(breaks = seq(from = 0, to = 400, by = 50)) + 
        scale_y_continuous(breaks = seq(from = 0, to = 20, by = 5)) +
        scale_color_manual(values = c("#8999b0","#738148","#7c5d2d","#447aab")) +
        transition_time(as.integer(year), range = c(2012L, 2020L))

animate(q1_plot, duration = 18, width = 8, height = 5, units = "in", res = 300)
```

---
# Question 1
```{r question 1 plot 2, echo = FALSE, warning = FALSE, message = FALSE}
parks_regions <- parks_q1 %>%
  mutate(region = case_when(
         city %in% c("Boston", "Long Beach", "New York", "Philadelphia") ~
           "Northeast",
         city %in% c("Atlanta", "Baltimore", "Jacksonville", "Louisville",
                    "Memphis", "Nashville", "Virginia Beach") ~ "Southeast",
         city %in% c("Chicago", "Columbus", "Detroit", "Kansas City",
                     "Milwaukee") ~ "Midwest",
         city %in% c("Albuquerque", "Austin", "Dallas", "El Paso",
                     "Fort Worth", "Houston", "Mesa", "Oklahoma City",
                     "Phoenix", "San Antonio", "Tucson") ~ "Southwest",
         city %in% c("Denver", "Fresno", "Las Vegas", "Los Angeles",
                     "Portland", "Sacramento", "San Diego", "San Francisco",
                     "San Jose", "Seattle") ~ "West"))

#group by region and year, then summarize mean of spending per resident and median park size
parks_regions <- parks_regions %>%
  group_by(region, year) %>%
  summarize(mean_spend = mean(spend_per_resident),
            mean_med_size = mean(med_park_size_data))

#create a simplified data set to annotate
regions <- parks_regions %>%
  count(year, region, mean_spend) %>%
  filter(year == 2019)

### plot of mean spending per resident with respect to mean of median park size over time

ggplot(data = parks_regions,
       aes(x = year, y = mean_spend, group = region)) +
  geom_line(aes(size = mean_med_size, color = region), lineend = "round",
            show.legend = FALSE) +
  geom_text_repel(data = regions,
                  aes(label = region, color = region),
                  show.legend = FALSE,
                  nudge_y = 9,
                  nudge_x = -1.75,
                  hjust = -.5) +
  scale_color_manual(values = c("#738148",
                                "#bc8a31",
                                "#3b5c75",
                                "#4f3e23",
                                "#8999b0")) +
  scale_y_continuous(breaks = seq(from = 0, to = 180, by = 20)) +
  labs(title = "Mean Spending per Resident Over Time with Respect to Mean of Median Park Size",
         subtitle = "by US Region",
         x = "Year",
         y = "Mean Spending Per Resident (in USD)") +
  theme(text = element_text(size = 18)) + 
  theme_minimal() 
<<<<<<< HEAD
=======

>>>>>>> 78aad82d45b9294fe088f7284c1fad39459430b4
```

---

class: inverse, middle, center

# Question 2

---
### How many amenities do parks with the top 10 and bottom 10 rankings in 2020 have and how does this vary based on what proportion of the top 10 and bottom 10 cities’ land is parkland in 2020?


--

- Focus in on Top 10 and Bottom 10 ranked cities in 2020

--

  - Top 10: Minneapolis, DC, St. Paul, Arlington, Cincinnati, Portland, Irvine, San Francisco, Boston, Chicago
  - Bottom 10: Memphis, Irving, Hialeah, Baton Rouge, Fresno, Lubbock, Forth Worth, Charlotte, Mesa, Oklahoma City

--

- Metrics of interest: percent of city that is parkland, geographic location, type of amenities and their numbers per residents

--

- Types of plots:
    - geographic map of the United States
    - stacked bar plot of amenities

---

# Question 2

```{r plots-2-1, echo = FALSE, warning = FALSE}
### data wrangling- 1st plot

#top/bottom 10 cities
parks_2020 <- parks %>%
  filter(year == 2020,
         rank <= 10 | rank >= 88)

#matching cities dataframe with parks dataframe
cities <- cities %>%
  filter(state != "Maine") %>%
  mutate(city = case_when(city == "Washington" ~ "Washington, D.C.",
                          city == "Charlotte" ~ "Charlotte/Mecklenburg County",
                          TRUE ~ city)) %>%
  select(city, latitude, longitude) %>%
  rbind(tibble(city = c("Arlington, Virginia"),
               latitude = c(38.8816),
               longitude = c(-77.0910)))

#merging cities and parks data frames
parks_2020_coords <- left_join(parks_2020, cities, by = "city")

#creating an indicator variable for rank
parks_2020_coords <- parks_2020_coords %>%
  mutate(rank_div = ifelse(rank <= 10, "top", "bottom"))

#dodging overlapping points
parks_2020_coords <- parks_2020_coords %>%
  mutate(longitude = case_when(rank == 1 ~ -93.5,
                              rank == 3 ~ -92.6,
                              rank == 2 ~ -76.6,
                              rank == 4 ~ -77.5,
                              rank == 89 ~ -96.7,
                              rank == 94 ~ -97.5,
                              TRUE ~ longitude),
         updown = ifelse(rank %in% c(3, 89, 2), "down", "up"))

### plot of top/bottom 10 cities scaled by % of parkland

ggplot() +
  geom_polygon(data = map_data("state"), aes(x = long, y = lat, group = group),
               fill = "white", color = "gray60") +
  geom_point(data = parks_2020_coords,
             aes(x = longitude, y = latitude, color = rank_div,
                 size = as.numeric(str_extract(park_pct_city_data,"\\d+"))/100)) +
  geom_text(data = parks_2020_coords %>% filter(updown == "up"),
            aes(x = longitude, y = latitude, label = paste0("#",rank)),
            size = 3.5, vjust = -.8, family = "bold") +
  geom_text(data = parks_2020_coords %>% filter(updown == "down"),
            aes(x = longitude, y = latitude, label = paste0("#",rank)),
            size = 3.5, vjust = 1.8, family = "bold") +
  scale_size_continuous(labels = scales::percent) +
  scale_color_manual(values = c("#bc8a31", "#315d1b")) +
  labs(x = NULL, y = NULL, size = "% of city that\nis parkland",
       title = "Top and bottom 10 city rankings of parks",
       subtitle = "scaled by % of city that is parkland") +
  coord_map() +
  theme_void() +
  guides(color = "none") +
  theme(legend.position = c(.92,.3),
        plot.title = element_text(hjust = 0.1),
        plot.subtitle = element_text(hjust = 0.1))
```

---

# Question 2

```{r plots-2-2, echo = FALSE, warning = FALSE}
### data wrangling-plot2

#creating a total amenities variable
parks_2020_coords <- parks_2020_coords %>%
  mutate(total_amenities = playground_data + restroom_data + basketball_data)

#creating a long dataset for amenities, adding rank to city name for plot, and
#shortening long city names
parks_amenities <- parks_2020_coords %>%
  pivot_longer(cols = c(playground_data, restroom_data, basketball_data),
               names_to = "amenity", values_to = "value") %>%
  mutate(city = ifelse(city == "Charlotte/Mecklenburg County", "Charlotte", city),
         city_n = paste0("#", rank, " ", city))

### plot of amenities

ggplot(data = parks_amenities, mapping = aes(y = reorder(city_n, -rank))) +
  geom_bar(stat = "identity", mapping = aes(x = value, fill = amenity)) +
  geom_hline(yintercept = 10.5, linetype = "dashed", color = "#322718") +
  guides(fill = guide_legend(reverse = TRUE)) +
  labs(title = "Top and bottom 10 city rankings by amenities",
       x = "Total Amenities per 10K  Residents", y = NULL, fill = "Amenities") +
  scale_fill_manual(values = c("#bc8a31", "#738148", "#3b5c75"),
                    labels = c("Basketball Courts", "Playgrounds", "Restrooms")) +
  theme_minimal() +
  theme(legend.position = c(0.8,0.2))
```

---

class: inverse, middle, center

# Thank You + Questions


